---
pagetitle: "CO deck example"
output: deckhand::co_deck
---

``` {r setup, include=FALSE}
library(ggplot2)
library(deckhand)

knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)

test_plot <- ggplot(
   palmerpenguins::penguins, 
   aes(x = flipper_length_mm, y = bill_length_mm)) + 
  geom_point(
    aes(colour = species, shape = species),
    size = 3, alpha = 0.8) +
  geom_smooth(aes(color = species),
              method = "lm", formula = y ~ x, se = FALSE) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Penguin size by species",
    x = "Flipper length (mm)",
    y = "Bill length (mm)"
  ) +
  theme_co_report()

```

::::{.grid-page .front-cover}

:::{.security-marking}
SECURITY-MARKING
:::

:::{.logo}
`r deckhand::co_logo()`
:::

:::{.title}
Report title
:::

:::{.panel}
:::

:::{.subtitle}
Report subtitle
:::

:::{.date}
DD Month YYYY<br />
Analysis & Insight, Cabinet Office
:::

::::

::::{.grid-page .single-content}

:::{.title}
{pagedown} reports for Cabinet Office
:::

:::{.maintext}
The `{deckhand}` package is a wrapper around the [`{pagedown}`](https://pagedown.rbind.io/) package to enable you to make reports in Rmarkdown that align (broadly) with the Cabinet Office's corporate identity guidelines.

A page is created in your Rmarkdown by creating a grid-page entity `::::{.grid-page .page-type}` where `.page-type` is the layout design you want to use (remember to close off with four colons `::::` before your next page).

You can then include page elements such as `:::{.title}`, `:::{.maintext}` and `:::{.content1}`, closing each element with `:::`. See below for an example.

The following pages show the supported layouts. You can also construct custom layouts based on the underlying grid system (explained at the end).

:::

:::{.content1}

````
::::{.grid-page .two-col}

:::{.title}
Page title
:::

:::{.maintext}
Main commentary text for the page
:::

:::{.content1}
```r
my_r_code
```
:::

:::{.content2}
```r
more_r_code
```
:::

::::
````

:::

::::

::::{.grid-page .divider}

:::{.panel}
:::

:::{.title .outline}
Divider title
:::

:::{.maintext .outline}
Use `.divider` to use this page layout. This is the `.maintext` element.
:::

::::

::::{.grid-page .single-content}

:::{.title .outline}
Single Content
:::

:::{.maintext .outline}
Use `.single-content-wide` to use this page layout. This is the `.maintext` element.
:::

:::{.secondarytext .outline}
This is the `.secondarytext` element.
:::

:::{.content1 .outline}
```{r 1c, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 8, height = 8)
```
:::

::::

::::{.grid-page .single-content-wide}

:::{.title .outline}
Single Content (wide)
:::

:::{.maintext .outline}
Use `.single-content-wide` to use this page layout. This is the `.maintext`.
:::


:::{.content1 .outline}
```{r 1w, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 12, height = 8)
```
:::

::::

::::{.grid-page .single-content-high}

:::{.title .outline}
Single Content (high)
:::

:::{.maintext .outline}
Use `.single-content-high` to use this page layout. This is the `.maintext`.
:::


:::{.content1 .outline}
```{r 1h, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 8, height = 12)
```
:::

::::

::::{.grid-page .two-col}

:::{.title .outline}
Two column
:::

:::{.maintext .outline}
Use `.two-col` to use this page layout. This is the `.maintext`.
:::

:::{.secondarytext .outline}
This is the `.secondarytext` element.
:::

:::{.content1 .outline}
```{r 2c_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 4, height = 8)
```
:::

:::{.content2 .outline}
```{r 2c_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 4, height = 8)
```
:::

::::

::::{.grid-page .two-col-wide}

:::{.title .outline}
Two column (wide)
:::

:::{.maintext .outline}
Use `.two-col-wide` to use this page layout. This is the `.maintext`.
:::

:::{.content1 .outline}
```{r 2w_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 6, height = 8)
```
:::

:::{.content2 .outline}
```{r 2w_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 6, height = 8)
```
:::

::::

::::{.grid-page .two-row}

:::{.title .outline}
Two row
:::

:::{.maintext .outline}
Use `.two-row` to use this page layout. This is the `.maintext`.
:::

:::{.secondarytext .outline}
This is the `.secondarytext` element.
:::

:::{.content1 .outline}
```{r 2r_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 8, height = 4)
```
:::

:::{.content2 .outline}
```{r 2r_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 8, height = 4)
```
:::

::::

::::{.grid-page .two-row-high}

:::{.title .outline}
Two row (high)
:::

:::{.maintext .outline}
Use `.two-row-high` to use this page layout. This is the `.maintext`.
:::

:::{.content1 .outline}
```{r 2h_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 8, height = 6)
```
:::

:::{.content2 .outline}
```{r 2h_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 8, height = 6)
```
:::

::::

::::{.grid-page .three-col}

:::{.title .outline}
Three column
:::

:::{.maintext .outline}
Use `.three-col` to use this page layout. This is the `.maintext`.
:::


:::{.content1 .outline}
```{r 3c_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(test_plot, width = 4, height = 8)
```
:::

:::{.content2 .outline}
```{r 3c_2, echo=FALSE, warning=FALSE, message=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 4, height = 8)
```
:::

:::{.content3 .outline}
```{r 3c_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 4, height = 8)
```
:::

::::

::::{.grid-page .three-row}

:::{.title .outline}
Three row
:::

:::{.maintext .outline}
Use `.three-row` to use this page layout. This is the `.maintext`.
:::


:::{.content1 .outline}
```{r 3r_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(test_plot, width = 8, height = 4)
```
:::

:::{.content2 .outline}
```{r 3r_2, echo=FALSE, warning=FALSE, message=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 8, height = 4)
```
:::

:::{.content3 .outline}
```{r 3r_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 8, height = 4)
```
:::

::::

::::{.grid-page .three-row-small}

:::{.title .outline}
Three row (small)
:::

:::{.maintext .outline}
Use `.three-row-small` to use this page layout. This is the `.maintext`.
:::

:::{.secondarytext .outline}
This is the `.secondarytext` element.
:::

:::{.content1 .outline}
```{r 3s_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content2 .outline}
```{r 3s_2, echo=FALSE, warning=FALSE, message=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content3 .outline}
```{r 3s_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 4, height = 4)
```
:::

::::

::::{.grid-page .four-content}

:::{.title .outline}
Four box content
:::

:::{.maintext .outline}
Use `.four-content` to use this page layout. This is the `.maintext`.
:::

:::{.secondarytext .outline}
This is the `.secondarytext` element.
:::

:::{.content1 .outline}
```{r 4b_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content2 .outline}
```{r 4b_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content3 .outline}
```{r 4b_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content4 .outline}
```{r 4b_4, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content4")

insert_svg(p, width = 4, height = 4)
```
:::

::::

::::{.grid-page .four-content-wide}

:::{.title .outline}
Four box content (wide)
:::

:::{.maintext .outline}
Use `.four-content-wide` to use this page layout. This is the `.maintext`.
:::

:::{.content1 .outline}
```{r 4w_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 6, height = 4)
```
:::

:::{.content2 .outline}
```{r 4w_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 6, height = 4)
```
:::

:::{.content3 .outline}
```{r 4w_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 6, height = 4)
```
:::

:::{.content4 .outline}
```{r 4w_4, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content4")

insert_svg(p, width = 6, height = 4)
```
:::

::::

::::{.grid-page .four-content-high}

:::{.title .outline}
Four box content (high)
:::

:::{.maintext .outline}
Use `.four-content-high` to use this page layout. This is the `.maintext`.
:::

:::{.content1 .outline}
```{r 4h_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 4, height = 6)
```
:::

:::{.content2 .outline}
```{r 4h_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 4, height = 6)
```
:::

:::{.content3 .outline}
```{r 4h_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 4, height = 6)
```
:::

:::{.content4 .outline}
```{r 4h_4, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content4")

insert_svg(p, width = 4, height = 6)
```
:::

::::

::::{.grid-page .four-col}

:::{.title .outline}
Four column
:::

:::{.maintext .outline}
Use `.four-col` to use this page layout. This is the `.maintext`.
:::

:::{.content1 .outline}
```{r 4c_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 3, height = 8)
```
:::

:::{.content2 .outline}
```{r 4c_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 3, height = 8)
```
:::

:::{.content3 .outline}
```{r 4c_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 3, height = 8)
```
:::

:::{.content4 .outline}
```{r 4c_4, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content4")

insert_svg(p, width = 3, height = 8)
```
:::

::::

::::{.grid-page .six-content-wide}

:::{.title .outline}
Six content (wide)
:::

:::{.maintext .outline}
Use `.six-content-wide` to use this page layout. This is the `.maintext`.
:::

:::{.content1 .outline}
```{r 6w_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content2 .outline}
```{r 6w_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content3 .outline}
```{r 6w_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content4 .outline}
```{r 6w_4, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content4")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content5 .outline}
```{r 6w_5, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content5")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content6 .outline}
```{r 6w_6, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content6")

insert_svg(p, width = 4, height = 4)
```
:::

::::

::::{.grid-page .six-content-high}

:::{.title .outline}
Six content (high)
:::

:::{.maintext .outline}
Use `.six-content-high` to use this page layout. This is the `.maintext`.
:::

:::{.content1 .outline}
```{r 6h_1, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content1")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content2 .outline}
```{r 6h_2, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content2")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content3 .outline}
```{r 6h_3, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content3")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content4 .outline}
```{r 6h_4, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content4")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content5 .outline}
```{r 6h_5, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content5")

insert_svg(p, width = 4, height = 4)
```
:::

:::{.content6 .outline}
```{r 6h_6, echo=FALSE, warning=FALSE}
p <- test_plot +
  labs(title = ".content6")

insert_svg(p, width = 4, height = 4)
```
:::

::::

::::{.grid-page .divider}

:::{.panel}
:::

:::{.title .outline}
Custom layouts
:::

::::

::::{.grid-page .demo}

:::{.title .outline}
Custom layouts
:::

:::{.start-row-1 .end-row-4 .start-column-5 .end-column-12 .maintext .outline}
The report template is built on a 12x12 grid. In addition to the set layouts shown before, bespoke layouts can also be constructed using the custom css classes. These are in the format `.start-row-X` and `.end-column-Y`, allowing you to specify the start and end rows and columns for a content box. Bear in mind the `.title` element is a 4x4 box starting in the upper left corner.

A standard `.maintext` element (such as this) occupies rows 1-4 and columns 5-12, to replicate this on a custom page use the classes `.start-row-1 .end-row-4 .start-column-5 .end-column-12 .maintext`, ensure you also include the `.maintext` class so that text flows across two columns.
:::

:::{.start-row-7 .end-row-9 .start-column-1 .end-column-3 .outline}
`.start-row-7 .end-row-9 .start-column-1 .end-column-3`
:::

:::{.start-row-6 .end-row-10 .start-column-6 .end-column-11 .outline}
`.start-row-6 .end-row-10 .start-column-6 .end-column-11`
:::

::::

::::{.grid-page .demo}

```{r 12grid, echo=FALSE, warning=FALSE}

div_maker <- function(x) {
  col <- x %% 12 + 1
  row <- x %/% 12 + 1
  htmltools::div(
    paste0("r", row, "c", col),
    style = paste0("grid-column: ", col, " / span 1; ",
                   "grid-row: ", row, " / span 1; ",
                   "background-color: var(--co-secondary-grey)")
  )
}

htmltools::tagList(purrr::map(.x = 0:143, .f = div_maker))

```

::::
